---
title: "Reproducible Research Project 1"
author: "Magnolia"
date: "28 October 2016"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Loading the necessary packages
```{r,echo=FALSE,results="hide"}
library(knitr)
library(dplyr)
library(lubridate)
library(ggplot2)
```
```{r,echo=FALSE,results="hide"}
setwd("C:/Users/Rudolph/Documents/Reproducible Research Project 1/Peer Assessment1")
```

##Loading and preprocessing the data

###Reading the data

```{r}
data <- read.csv("activity.csv", colClass=c('integer', 'Date', 'integer'))
```

###Tyding the data

Changing the date into dateformat using lubridate:
```{r}
data$date <- ymd(data$date)
```

##What is mean total number of steps taken per day?


###Methodology and Result

1.Calculating the total number of steps per day using dplyr and group by date:
```{r}
steps <- data %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print
```

2.Using ggplot for making the histogram:
```{r}
ggplot(steps, aes(x = steps)) +
  geom_histogram(fill = "purple", binwidth = 1000) +
  labs(title = "Histogram of Steps per day", x = "Steps per day", y = "Frequency")
```

3.Calculating the mean and median of the total number of steps taken per day:
```{r}
mean_steps <- mean(steps$steps, na.rm = TRUE)
median_steps <- median(steps$steps, na.rm = TRUE)
```
```{r}
mean_steps
```
```{r}
median_steps
```

Mean steps are 10766 and median steps are 10765.


##What is the average daily activity pattern?

###Methodology and Result

1. Calculating the average number of steps taken in each 5-minute interval per day using dplyr and group by interval:

```{r}
interval <- data %>%
  filter(!is.na(steps)) %>%
  group_by(interval) %>%
  summarize(steps = mean(steps))
```

Using ggplot for making the time series of the 5-minute interval and average steps taken:

```{r}
ggplot(interval, aes(x=interval, y=steps)) +
  geom_line(color = "purple")
```

2. Using which.max() to find out the maximum steps, on average, across all the days:

```{r}
interval[which.max(interval$steps),]
```

The interval 835 has, on average, the highest count of steps, with 206 steps.


##Imputing missing values

###Methodology and Result

Summarizing all the missing values:

```{r}
sum(is.na(data$steps))
```

Missing values are 2304.

1. Taking the approach to fill in a missing NA with the average number of steps in the same 5-min interval.
2. Creating a new dataset as the original and fill in the missing values with the average number of steps per 5-minute interval:

```{r}
data_full <- data
nas <- is.na(data_full$steps)
avg_interval <- tapply(data_full$steps, data_full$interval, mean, na.rm=TRUE, simplify=TRUE)
data_full$steps[nas] <- avg_interval[as.character(data_full$interval[nas])]
```

Checking that there are no missing values:

```{r}
sum(is.na(data_full$steps))
```

No more missing values.

4. Calculating the number of steps taken in each 5-minute interval per day using dplyr and group by interval. Using ggplot for making the histogram:

```{r}
steps_full <- data_full %>%
  filter(!is.na(steps)) %>%
  group_by(date) %>%
  summarize(steps = sum(steps)) %>%
  print
```

```{r}
ggplot(steps_full, aes(x = steps)) +
  geom_histogram(fill = "purple", binwidth = 1000) +
  labs(title = "Histogram of Steps per day, including missing values", x = "Steps per day", y = "Frequency")
```


Calculating the mean and median steps with the filled in values:

```{r}
mean_steps_full <- mean(steps_full$steps, na.rm = TRUE)
median_steps_full <- median(steps_full$steps, na.rm = TRUE)
```

```{r}
mean_steps_full
```

```{r}
median_steps_full
```

The impact of imputing missing data with the average number of steps in the same 5-min interval is that both the mean and the median are equal to the same value: 10766.

##Are there differences in activity patterns between weekdays and weekends?

###Methodology and Result

Using dplyr and mutate to create a new column, weektype, and apply whether the day is weekend or weekday:

```{r}
data_full <- mutate(data_full, weektype = ifelse(weekdays(data_full$date) == "Saturday" | weekdays(data_full$date) == "Sunday", "weekend", "weekday"))
data_full$weektype <- as.factor(data_full$weektype)
head(data_full)
```

2. Calculating the average steps in the 5-minute interval and use ggplot for making the time series of the 5-minute interval for weekday and weekend, and compare the average steps:

```{r}
interval_full <- data_full %>%
  group_by(interval, weektype) %>%
  summarise(steps = mean(steps))
s <- ggplot(interval_full, aes(x=interval, y=steps, color = weektype)) +
  geom_line() +
  facet_wrap(~weektype, ncol = 1, nrow=2)
print(s)
```

From the two plots it seems that the test object is more active earlier in the day during weekdays compared to weekends, but more active throughout the weekends compared with weekdays (probably because the oject is working during the weekdays, hence moving less during the day).
